\thetitle

### 2021-02-23 writing/rational.tex
small fix
organize sections/titles/labels in paper
simple 'we' cleanup

\section{How to Test the Rational Programmer} 
\label{sec:rational}

@@

Based on this idea, migratory paths offer a concise setting for
evaluating the effectiveness of a rational programmer for dealing with the
interesting debugging scenarios  from section ~\ref{sec:mutate}.
If we group debugging scenarios by mutation, then we obtain a set 
of programs that all contain the same mistake but differ in their type
annotations.  In the setting of migratory typing,\footnote{The method can be adapted to
so-called true gradual typing systems~\cite{svcb-snapl-2015} in a
straightforward manner.} each scenario differs from another in which
components are typed. Formally, a program $\system$ is a set of
components $\set{\component}$, and a scenario $\conf$ is the typed subset of
$\set{\component}$ . Hence, the scenarios of $\system$ are ordered by the
subset relation and, naturally,  form a lattice $\lattice{\system}$ with
$2^{\size{\system}}$ elements, like
the one introduced by ~\citet{tfgnvf-popl-2016} for performance
evaluations. The bottom scenario of
$\lattice{\system}$ is always $\emptyset$ and the top one is $\system$
itself. In between are all the mix-typed scenarios.

With these definitions, the actions of the rational programmer map to an
ascending chain---dubbed a \emph{trail}---in $\lattice{\system}$.  The
starting scenario $\conf_0$ of a trail---also referred to as the
\emph{root} (of the trail)---is the initial debugging scenario.  Herein we
consider only roots that are interesting debugging scenarios as described
in~\ref{sec:mutate} --- we expand on this point further on.  Intuitively,
starting from such a root, the rational programmer  steps along a trail,
and at each step, reduces the current scenario to an ``easier'' next
scenario to debug.  The easiest scenarios are those that the type checker
rejects outright, because a type error pinpoints an impedance mismatch
between the broken code of one component and its typed interface to
others. At this point, the programmer has identified the broken component.
If the scenario type-checks, the programmers runs the program until it
raises a run time error.  The rational programmer uses the error's blame
assignment\footnote{Blame can result either directly from a run time type
check or, in the absence of blame, from an exception that we track back to
the component that causes it with the help of the stack trace.} to decide
which component(s) to equip with types. This choice constructs scenario
$\conf_{i+1}$ from $\conf_{i}$ and implies the creation of trail.

@@

The next three subsections specialize the trail-scenario idea to the
various semantics of gradual typing. Once this machinery is in place,
we use the fourth subsection to state the precise experimental
research questions that correspond to the high-level question on the
first page and to explain our experimental process. The final
subsection extends the development with a discussion of programmer effort.

>>>>

@@

%% MF: I don't think we should honor the nonsense of ``true gt.''  We
% are it as much as anyone!
% 
% \footnote{The method can be adapted to so-called true gradual typing
% systems~\cite{svcb-snapl-2015} in a straightforward manner.}

Based on this idea, migratory paths offer a concise setting for
evaluating the effectiveness of a rational programmer for dealing with
the interesting debugging scenarios from section ~\ref{sec:mutate}.
Grouping debugging scenarios by mutation yields a set of programs that
all contain the same mistake but differ in their type annotations.  In
the setting of migratory typing, each scenario differs from another in
which components are typed. Formally, a program $\system$ is a set of
components $\set{\component}$, and a scenario $\conf$ is the typed
subset of $\set{\component}$ . Hence, the scenarios of $\system$ are
ordered by the subset relation and, naturally, form a lattice
$\lattice{\system}$ with $2^{\size{\system}}$
elements.\footnote{\citet{tfgnvf-popl-2016} use this lattice for
performance evaluations.} The bottom scenario of $\lattice{\system}$
is always $\emptyset$ and the top one is $\system$ itself. In between
are all the mixed-typed scenarios.

In this context, the actions of the rational programmer map to an
ascending chain---dubbed a \emph{trail}---in $\lattice{\system}$.  The
starting scenario $\conf_0$ of a trail---also referred to as the
\emph{root} (of the trail)---is the initial debugging scenario; only
interesting (as described in~\ref{sec:mutate}) debugging scenarios
serve as roots.  Intuitively, starting from such a root, the rational
programmer steps along a trail, and at each step, reduces the current
scenario to an ``easier'' next scenario to debug.  The easiest
scenarios are those that the type checker rejects outright, because a
type error pinpoints an impedance mismatch between the broken code of
one component and its typed interface to others. At this point, the
programmer has identified the broken component.  If the scenario
type-checks, the programmers runs the program until it raises a run
time error.  The rational programmer uses the error's blame
assignment\footnote{Blame can result either directly from a run-time
check or, in the absence of blame, from an exception that can be
traced back to the component that causes it with the help of the stack
trace.} to decide which component(s) to equip with types. This choice
constructs scenario $\conf_{i+1}$ from $\conf_{i}$ and thus creates a trail.

@@

The first three subsections specialize the trail-scenario idea to the
three semantics of gradual typing. Once this machinery is in place,
the fourth subsection presents a reformulation of the informal
question on page~1 as precise experimental research questions. The
final subsection adds a notion of programmer effort to the
experimental setup.
